% ARKHEION AGI 2.0 - Paper 30: Multi-Personality System
% Jhonatan Vieira Feitosa | Manaus, Amazonas, Brazil
% February 2026

\documentclass[11pt,twocolumn]{article}

% Encoding and fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

% Layout
\usepackage[margin=0.75in]{geometry}
\usepackage{fancyhdr}

% Mathematics
\usepackage{amsmath,amssymb}

% Graphics and colors
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,shapes,positioning}

% Tables
\usepackage{booktabs}

% Code listings
\usepackage{listings}

% Hyperlinks
\usepackage{hyperref}

% ==================== COLORS ====================
\definecolor{arkblue}{RGB}{0,102,204}
\definecolor{arkpurple}{RGB}{102,51,153}
\definecolor{arkgreen}{RGB}{0,153,76}
\definecolor{arkgold}{RGB}{218,165,32}

% ==================== LISTINGS ====================
\lstset{
    basicstyle=\ttfamily\scriptsize,
    breaklines=true,
    breakatwhitespace=true,
    postbreak=\mbox{\textcolor{gray}{$\hookrightarrow$}\space},
    columns=flexible,
    keepspaces=true,
    showstringspaces=false,
    numbers=none,
    backgroundcolor=\color{gray!5},
    frame=single,
    rulecolor=\color{gray!30}
}

% ==================== HEADER/FOOTER ====================
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{arkblue}{ARKHEION AGI 2.0}}
\fancyhead[R]{\small Paper 30: Multi-Personality}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ==================== HYPERREF ====================
\hypersetup{
    colorlinks=true,
    linkcolor=arkblue,
    urlcolor=arkpurple,
    citecolor=arkgreen
}

% ==================== TITLE ====================
\title{
    \vspace{-1.5cm}
    {\Large\textbf{Multi-Personality Architecture}}\\[0.3em]
    {\large Adaptive Personas in AGI}\\[0.2em]
    {\normalsize ARKHEION AGI 2.0 --- Paper 30}
}

\author{Jhonatan Vieira Feitosa\
Independent Researcher\
\texttt{ooriginador@gmail.com}\
Manaus, Amazonas, Brazil}

\date{February 2026}

\begin{document}

\maketitle

% ==================== ABSTRACT ====================
\begin{abstract}
\noindent
This paper presents \textbf{Multi-Personality Architecture}, a system enabling ARKHEION AGI 2.0 to adopt context-appropriate personas. The architecture includes a \textbf{Personality Engine}, \textbf{Emotion Simulator}, and \textbf{Context Analyzer} that dynamically switch between specialized personalities (Scientist, Artist, Engineer, Philosopher, Therapist). Using $\phi$-enhanced optimization and Big Five personality traits, the system achieves \textbf{persona consistency of 94\%} and \textbf{context-switch latency under 50ms}.

\vspace{0.5em}
\noindent\textbf{Keywords:} personality, personas, emotion simulation, Big Five, context switching, AGI
\end{abstract}

% ==================== EPISTEMOLOGICAL NOTE ====================
\section*{Epistemological Note}
\textit{This paper distinguishes between \textbf{heuristic} concepts and \textbf{empirical} results:}

\begin{center}
\footnotesize
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Heuristic} & \textbf{Empirical} \\
\midrule
``Personality'' & Consistency: 94\% \\
``Emotion simulation'' & Switch latency: <50ms \\
``Psychological model'' & 617 LOC main file \\
\bottomrule
\end{tabular}
\end{center}

% ==================== INTRODUCTION ====================
\section{Introduction}

Effective AGI must adapt its communication style to context. A system explaining quantum physics to a child differs from one presenting to experts. ARKHEION's Multi-Personality System enables:

\begin{itemize}
    \item \textbf{Context-aware persona selection}
    \item \textbf{Emotion-modulated responses}
    \item \textbf{Specialized expertise activation}
    \item \textbf{Smooth personality transitions}
\end{itemize}

% ==================== ARCHITECTURE ====================
\section{System Architecture}

\subsection{Core Components}

\begin{center}
\begin{tikzpicture}[
    box/.style={rectangle, draw=arkblue, fill=arkblue!10, rounded corners, minimum width=2.2cm, minimum height=0.7cm, font=\scriptsize},
    arrow/.style={-{Stealth}, thick}
]
    \node[box] (context) at (0,2) {Context Analyzer};
    \node[box] (engine) at (2.5,2) {Personality Engine};
    \node[box] (emotion) at (5,2) {Emotion Simulator};
    \node[box] (personas) at (2.5,0.5) {Persona Library};

    \draw[arrow] (context) -- (engine);
    \draw[arrow] (engine) -- (emotion);
    \draw[arrow] (personas) -- (engine);
\end{tikzpicture}
\end{center}

\subsection{System Metrics}

\begin{lstlisting}[language=Python]
@dataclass
class ARKHEIONSystemMetrics:
    total_interactions: int = 0
    successful_switches: int = 0
    average_response_time: float = 0.0
    user_satisfaction: float = 0.0
    system_uptime: float = 0.0
    phi_optimization_level: float = 0.0
\end{lstlisting}

% ==================== PERSONALITY ENGINE ====================
\section{Personality Engine}

\subsection{Big Five Traits}

Each personality is defined by OCEAN traits:

\begin{center}
\footnotesize
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{Persona} & \textbf{O} & \textbf{C} & \textbf{E} & \textbf{A} & \textbf{N} \\
\midrule
Scientist & 0.9 & 0.95 & 0.5 & 0.6 & 0.3 \\
Artist & 0.95 & 0.4 & 0.8 & 0.7 & 0.5 \\
Engineer & 0.7 & 0.9 & 0.4 & 0.5 & 0.2 \\
Philosopher & 0.95 & 0.6 & 0.6 & 0.7 & 0.4 \\
Therapist & 0.6 & 0.7 & 0.8 & 0.95 & 0.3 \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Legend:} O=Openness, C=Conscientiousness, E=Extraversion, A=Agreeableness, N=Neuroticism

\textbf{Note:} Big Five personality trait values (e.g., Openness=0.9, Neuroticism=0.3) are manually assigned design parameters, not psychometrically validated measurements. They were chosen to produce distinguishable behavioral profiles for each persona.

\subsection{Transition Modes}

\begin{lstlisting}[language=Python]
class TransitionMode(Enum):
    INSTANT = "instant"      # Immediate switch
    GRADUAL = "gradual"      # Smooth blend
    CONTEXTUAL = "contextual" # Context-driven
\end{lstlisting}

% ==================== EMOTION SIMULATOR ====================
\section{Emotion Simulator}

\subsection{Emotional States}

\begin{center}
\footnotesize
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Emotion} & \textbf{Valence} & \textbf{Arousal} \\
\midrule
Joy & +0.8 & +0.6 \\
Curiosity & +0.5 & +0.7 \\
Calm & +0.4 & -0.3 \\
Concern & -0.2 & +0.4 \\
Focus & +0.3 & +0.2 \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Emotion Blending}

Emotions blend based on context:

\begin{equation}
E_{blend} = \sum_{i} w_i \cdot E_i, \quad \sum w_i = 1
\end{equation}

where weights $w_i$ come from context analysis.

% ==================== CONTEXT ANALYZER ====================
\section{Context Analyzer}

\subsection{Context Signals}

\begin{itemize}
    \item \textbf{Topic}: Scientific, artistic, technical
    \item \textbf{Formality}: Casual to academic
    \item \textbf{Expertise}: Beginner to expert
    \item \textbf{Emotional tone}: Supportive, neutral, challenging
\end{itemize}

\subsection{Persona Selection}

\begin{lstlisting}[language=Python]
def select_persona(self, context: dict) -> str:
    topic = context.get("topic", "general")
    expertise = context.get("expertise", 0.5)

    if topic in ["physics", "math", "biology"]:
        return "Scientist"
    elif topic in ["art", "music", "design"]:
        return "Artist"
    elif topic in ["engineering", "coding"]:
        return "Engineer"
    elif expertise > 0.8:
        return "Philosopher"
    else:
        return "Therapist"  # Default supportive
\end{lstlisting}

% ==================== PHI OPTIMIZATION ====================
\section{$\phi$-Enhanced Optimization}

Sacred geometry constants optimize transitions:

\begin{lstlisting}[language=Python]
PHI = 1.618033988749895
GOLDEN_ANGLE = 137.508

def phi_optimized_blend(self, traits_a, traits_b, t):
    """Blend traits using golden ratio timing."""
    phi_t = t ** (1/PHI)  # Non-linear easing
    return (1-phi_t) * traits_a + phi_t * traits_b
\end{lstlisting}

\textbf{Note:} The $\varphi$-based easing function ($t^{1/\varphi}$) produces a specific non-linear interpolation curve. This was chosen for aesthetic consistency with the project's sacred geometry theme; no comparison with standard easing functions (quadratic, cubic, sigmoid) was performed to evaluate whether this produces measurably better transitions.

% ==================== SPECIALIZED PERSONAS ====================
\section{Specialized Personas}

\subsection{Scientist Personality}

\begin{itemize}
    \item \textbf{Style}: Precise, evidence-based
    \item \textbf{Language}: Technical, citations
    \item \textbf{Approach}: Hypothesis-driven
\end{itemize}

\subsection{Artist Personality}

\begin{itemize}
    \item \textbf{Style}: Expressive, metaphorical
    \item \textbf{Language}: Evocative, imagery
    \item \textbf{Approach}: Creative exploration
\end{itemize}

\subsection{Therapist Personality}

\begin{itemize}
    \item \textbf{Style}: Empathetic, supportive
    \item \textbf{Language}: Warm, validating
    \item \textbf{Approach}: Active listening
\end{itemize}

% ==================== RESULTS ====================
\section{Experimental Results}

\subsection{Consistency Metrics}

\begin{center}
\footnotesize
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Persona} & \textbf{Consistency} & \textbf{Switch Time} \\
\midrule
Scientist & 96\% & 42ms \\
Artist & 92\% & 38ms \\
Engineer & 95\% & 45ms \\
Philosopher & 93\% & 48ms \\
Therapist & 94\% & 35ms \\
\midrule
\textbf{Average} & \textbf{94\%} & \textbf{42ms} \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Measurement note:} Persona consistency (94\%) is measured as the fraction of generated responses that match the active persona's Big Five trait direction (e.g., high-extraversion persona produces sociable responses) on a 100-response evaluation set judged by the system's own classifier. No human evaluation was performed.

\subsection{User Satisfaction}

\begin{center}
\footnotesize
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{Metric} & \textbf{Score} \\
\midrule
Response appropriateness & 4.2/5.0 \\
Personality coherence & 4.4/5.0 \\
Emotional resonance & 4.1/5.0 \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Important:} The satisfaction scores above (4.2/5.0, 4.4/5.0, 4.1/5.0) are \textbf{design targets}, not measured outcomes. No user study has been conducted. Validating these targets requires a controlled user study with defined participant recruitment, sample size, protocol, and (if applicable) IRB approval.

% ==================== IMPLEMENTATION ====================
\section{Implementation}

\begin{center}
\footnotesize
\begin{tabular}{@{}lr@{}}
\toprule
\textbf{File} & \textbf{Lines} \\
\midrule
arkheion\_multi\_personality\_system.py & 617 \\
unified\_integration\_system.py & 1,054 \\
unified\_integration\_demo.py & 865 \\
\midrule
\textbf{Total} & \textbf{2,536} \\
\bottomrule
\end{tabular}
\end{center}

% ==================== CONCLUSION ====================
\section{Conclusion}

Multi-Personality Architecture enables ARKHEION AGI 2.0 to adopt context-appropriate personas with high consistency and low latency. The Big Five trait model and $\phi$-optimized transitions create coherent, adaptive behavior.\footnote{Implementation update (Feb 2026): The multi-personality ecosystem has since expanded from the 2,536 SLOC across 3 files described here to approximately 30,500 SLOC across \texttt{src/advanced/personality/} and \texttt{src/arkheion/personalities/}, incorporating behavior adaptation, context awareness, advanced personality metrics, and domain-specific personality profiles (engineer, scientist, artist, therapist).}

\textbf{Future work}:
\begin{itemize}
    \item Learning new personas from interaction
    \item Cultural adaptation
    \item Long-term personality evolution
\end{itemize}

% ==================== REFERENCES ====================
\section*{References}

\begin{enumerate}
\footnotesize
    \item Goldberg, L.R. ``The structure of phenotypic personality traits.'' American Psychologist, 1993.
    \item Papers 14, 31 of ARKHEION AGI 2.0 series.
\end{enumerate}

\end{document}
