% ARKHEION AGI 2.0 - Paper 39: Ternary Gene Synthesis
% Synthetic Parameter Generation via Evolutionary Optimization
% Jhonatan Vieira Feitosa | Manaus, Amazonas, Brazil
% February 2026

\documentclass[11pt,twocolumn]{article}

% Encoding and fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

% Layout
\usepackage[margin=0.75in]{geometry}
\usepackage{fancyhdr}

% Mathematics
\usepackage{amsmath,amssymb}

% Graphics and colors
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,shapes,positioning}

% Tables
\usepackage{booktabs}

% Code listings
\usepackage{listings}

% Hyperlinks
\usepackage{hyperref}

% Float control
\usepackage{float}

% Plots
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

% ==================== COLORS ====================
\definecolor{arkblue}{RGB}{0,102,204}
\definecolor{arkpurple}{RGB}{102,51,153}
\definecolor{arkgreen}{RGB}{0,153,76}
\definecolor{arkorange}{RGB}{255,128,0}
\definecolor{arkgold}{RGB}{218,165,32}
\definecolor{arkred}{RGB}{204,51,51}

% ==================== LISTINGS ====================
\lstset{
    language=Python,
    basicstyle=\ttfamily\scriptsize,
    keywordstyle=\color{arkblue},
    stringstyle=\color{arkgreen},
    commentstyle=\color{gray}\itshape,
    breaklines=true,
    breakatwhitespace=true,
    postbreak=\mbox{\textcolor{gray}{$\hookrightarrow$}\space},
    columns=flexible,
    keepspaces=true,
    showstringspaces=false,
    numbers=none,
    backgroundcolor=\color{gray!5},
    frame=single,
    rulecolor=\color{gray!30}
}

% ==================== HEADER/FOOTER ====================
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textcolor{arkblue}{ARKHEION AGI 2.0}}
\fancyhead[R]{\small Paper 39: Gene Synthesis}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ==================== HYPERREF ====================
\hypersetup{
    colorlinks=true,
    linkcolor=arkblue,
    urlcolor=arkpurple,
    citecolor=arkgreen
}

% ==================== TITLE ====================
\title{
    \vspace{-1.5cm}
    {\Large\textbf{Ternary Gene Synthesis}}\\[0.3em]
    {\large Evolutionary Optimization of Discrete Neural Parameters}\\[0.2em]
    {\normalsize ARKHEION AGI 2.0 --- Paper 39}
}

\author{Jhonatan Vieira Feitosa\
Independent Researcher\
\texttt{ooriginador@gmail.com}\
Manaus, Amazonas, Brazil}

\date{February 2026}

\begin{document}

\maketitle

% ==================== ABSTRACT ====================
\begin{abstract}
\noindent
We present a complete pipeline for \textbf{synthetic gene generation} in ternary neural networks, where weights are constrained to $\{-1, 0, +1\}$. This discrete parameter space enables novel optimization strategies impossible in continuous-weight networks: statistical pattern extraction, genetic interpolation, motif-based construction, and multi-generational evolutionary search. Our system implements seven synthesis methods (interpolation, mutation, crossover, distribution sampling, motif composition, adversarial generation, and $\varphi$-optimized perturbation), a proxy fitness evaluator achieving \textbf{8,229 genes/second} on GPU, and a full evolutionary engine with adaptive mutation, diversity maintenance, and island-model parallelism. Empirical evaluation on 200-gene pools with GPT-2-scale dimensions shows best fitness of \textbf{0.9267} in single-pass synthesis (427.8\,ms) and \textbf{2.5419} after 30 evolutionary generations (0.83\,s). Population compression via HTCV4 achieves \textbf{9.0:1} ratios on evolved populations. Implementation: 2,536 LOC Python, 41/41 unit tests passing.

\vspace{0.5em}
\noindent\textbf{Keywords:} ternary quantization, neuroevolution, gene synthesis, discrete optimization, parameter-space search
\end{abstract}

% ==================== EPISTEMOLOGICAL NOTE ====================
\section*{Epistemological Note}

\textit{This paper explicitly distinguishes between \textbf{heuristic} concepts (metaphors guiding design) and \textbf{empirical} results (measurable outcomes).}

\begin{center}
\footnotesize
\begin{tabular}{@{}ll@{}}
\toprule
\textbf{Heuristic} & \textbf{Empirical} \\
\midrule
``Gene'', ``evolution'', ``fitness'' & Best fitness: 0.9267 / 2.5419 \\
``Mutation'', ``crossover'' & GPU throughput: 8,229 genes/s \\
Biological metaphors & HTCV4 compression: 9.0:1 \\
$\varphi$-optimization & Pipeline time: 427.8\,ms \\
\bottomrule
\end{tabular}
\end{center}

A ``gene'' in this context is a ternary weight vector $\mathbf{w} \in \{-1,0,+1\}^n$, not a biological entity. ``Evolution'' denotes iterative stochastic optimization, not biological natural selection.

% ==================== INTRODUCTION ====================
\section{Introduction}

Traditional neural network optimization operates in continuous parameter spaces via gradient descent. When weights are quantized to ternary values $\{-1, 0, +1\}$, the parameter space becomes \textbf{discrete and finite}: a network layer with $n$ parameters has exactly $3^n$ possible configurations. This discreteness, typically viewed as a constraint, enables a fundamentally different optimization paradigm.

We introduce \textbf{Gene Synthesis}---a pipeline that treats ternary weight vectors as ``genes'' subject to analysis, recombination, and evolutionary optimization. The key insight is that ternary parameters admit:

\begin{enumerate}
    \item \textbf{Exact statistical analysis}: Distribution, sparsity, and motif patterns can be computed directly without approximation.
    \item \textbf{Meaningful interpolation}: Between-gene mixing operates on discrete symbols, producing valid configurations without rounding artifacts.
    \item \textbf{Efficient proxy evaluation}: Structural properties (sparsity balance, sign symmetry, local coherence, entropy) serve as fast fitness proxies.
    \item \textbf{Lossless compression}: HTCV4 trit-packing stores populations at 9.0:1 ratios.
\end{enumerate}

Our system implements this as a four-stage pipeline: \textsc{Analyze} $\rightarrow$ \textsc{Generate} $\rightarrow$ \textsc{Evaluate} $\rightarrow$ \textsc{Inject}, with optional evolutionary multi-generation refinement.

% ==================== BACKGROUND ====================
\section{Background}

\subsection{Ternary Neural Networks}

Ternary Weight Networks (TWN)~\cite{li2016ternary} constrain weights to $\{-1, 0, +1\}$, reducing model size by $16\times$ versus FP32 and enabling bitwise operations for inference. The ARKHEION Nucleus extends this with \textbf{gene-level granularity}: each weight tensor in a transformer layer is treated as an independent gene with its own hash, metadata, and evolutionary history.

\subsection{Neuroevolution}

Neuroevolution applies evolutionary algorithms to neural network optimization~\cite{stanley2019designing}. Prior work focuses on continuous weights or architecture search. Our contribution applies evolution specifically to \textbf{discrete ternary parameters}, where the finite alphabet enables exact diversity calculation and lossless population storage.

\subsection{Golden Ratio Optimization}

The golden ratio $\varphi = 1.618033...$ provides optimal coverage of parameter spaces via the golden angle $\theta = 2\pi/\varphi^2 \approx 137.5°$~\cite{vogel1979better}. We use $\varphi$-spaced perturbations for maximally uniform parameter exploration.

% ==================== ARCHITECTURE ====================
\section{Architecture}

The Gene Synthesis system comprises seven classes organized into three tiers:

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    box/.style={draw, rounded corners, minimum width=2.4cm, minimum height=0.7cm, font=\scriptsize\ttfamily, fill=#1!15},
    arrow/.style={->, thick}
]
% Analysis tier
\node[box=arkblue] (analyzer) at (0,4) {GeneAnalyzer};
\node[box=arkblue] (diversity) at (3.2,4) {DiversityCalc};

% Generation tier
\node[box=arkpurple] (generator) at (0,2.7) {SyntheticGeneGen};
\node[box=arkpurple] (engine) at (3.2,2.7) {EvolutionEngine};

% Evaluation tier
\node[box=arkgreen] (evaluator) at (0,1.4) {FastEvaluator};
\node[box=arkgreen] (gpueval) at (3.2,1.4) {GPUBatchEval};

% Integration tier
\node[box=arkgold] (compressor) at (0,0.1) {PopCompressor};
\node[box=arkgold] (synth) at (3.2,0.1) {GeneSynthesizer};

% Arrows
\draw[arrow] (analyzer) -- (generator);
\draw[arrow] (generator) -- (evaluator);
\draw[arrow] (evaluator) -- (synth);
\draw[arrow] (diversity) -- (engine);
\draw[arrow] (engine) -- (gpueval);
\draw[arrow] (gpueval) -- (compressor);
\draw[arrow, dashed] (synth) -- (engine);
\end{tikzpicture}
\caption{Gene Synthesis architecture. Blue: analysis; purple: generation; green: evaluation; gold: integration.}
\end{figure}

\subsection{GeneAnalyzer}

Extracts statistical patterns from gene pools:

\begin{itemize}
    \item \textbf{Distribution}: Counts of $\{-1, 0, +1\}$ across all genes.
    \item \textbf{Sparsity}: Fraction of zero weights per gene.
    \item \textbf{Motifs}: Common subsequences of length $k=8$, extracted via sampling ($\leq 1000$ samples/gene).
    \item \textbf{Clusters}: Similarity-based grouping with threshold $\tau=0.8$.
\end{itemize}

For a pool of 200 genes with GPT-2-scale dimensions (768--4096), analysis produces: distribution $(-1: 29.0\%, 0: 42.0\%, +1: 29.1\%)$, sparsity $42.1\% \pm 1.2\%$, 20 motifs, and 50 clusters.

\subsection{Synthesis Methods}

Seven methods generate new genes from existing ones:

\begin{lstlisting}[title={Ternary Gene Interpolation}]
Input: Genes a, b in {-1,0,+1}^n, mixing alpha
r = copy(a)                # Start from parent A
for i where a[i] != b[i]:
    if rand() >= alpha:
        r[i] = b[i]        # Take from parent B
return r
\end{lstlisting}

\textbf{Mutation} applies three strategies vectorized via NumPy: sign flip ($-1 \leftrightarrow +1$), zero toggle ($0 \leftrightarrow \pm 1$), and full random replacement.

\textbf{$\varphi$-Optimization} distributes changes at golden-angle intervals:
\begin{equation}
    p_k = \left\lfloor \frac{k \cdot \theta}{2\pi} \cdot n \right\rfloor \bmod n, \quad \theta = \frac{2\pi}{\varphi^2}
\end{equation}
ensuring maximally uniform coverage of the parameter space. Weight cycling uses $(w + 2) \bmod 3 - 1$, which rotates $-1 \rightarrow 0 \rightarrow 1 \rightarrow -1$.

\subsection{Proxy Fitness Evaluation}

The \texttt{FastEvaluator} computes fitness without neural network forward passes, using four structural metrics combined with $\varphi$-weighted importance:

\begin{equation}
    F(\mathbf{w}) = \frac{\varphi \cdot S_\text{sp} + S_\text{bal} + \frac{S_\text{coh}}{\varphi} + \frac{S_\text{ent}}{\varphi^2}}{\varphi + 1 + \frac{1}{\varphi} + \frac{1}{\varphi^2}}
\end{equation}

\noindent\textbf{Weight Justification:} The $\varphi$-based weighting (sparsity$\times\varphi$, coherence$\times 1/\varphi$, entropy$\times 1/\varphi^2$) is an aesthetic design choice. No ablation comparing these weights to uniform or optimized alternatives has been conducted.

where:
\begin{align}
    S_\text{sp} &= 1 - 2 \cdot |\text{sparsity}(\mathbf{w}) - 0.35| \\
    S_\text{bal} &= 1 - 2 \cdot |P(w_i > 0 \mid w_i \neq 0) - 0.5| \\
    S_\text{coh} &= 1 - \frac{\overline{|w_{i+1} - w_i|}}{2} \\
    S_\text{ent} &= \frac{H(\mathbf{w})}{\log 3}
\end{align}

Large arrays ($>100$K elements) are evaluated on sampled subsets for $O(1)$ cost.

\subsection{Evolutionary Engine}

The \texttt{EvolutionaryEngine} implements multi-generational optimization with:

\begin{itemize}
    \item \textbf{Selection}: Tournament ($k=3$), roulette (fitness-proportionate), or rank-based with configurable selection pressure $s = \varphi$.
    \item \textbf{Adaptive mutation}: Rate increases $1.2\times$ during stagnation, decreases $0.9\times$ during progress. Range: $[0.01, 0.5]$.
    \item \textbf{Elite preservation}: Top $e=5$ genes survive unconditionally.
    \item \textbf{Diversity maintenance}: Rarity bonus $\delta$ for uncommon patterns:
    \begin{equation}
        F'(\mathbf{w}) = F(\mathbf{w}) + \delta \cdot \left(1 - \frac{\text{count}(\text{pat}(\mathbf{w}))}{\max_\text{count}}\right)
    \end{equation}

    \noindent\textbf{Fitness Range Note:} The base fitness $F(\mathbf{w}) \in [0, 1]$ (Eq.~above), but the rarity bonus $\delta$ adds up to 1.8 for unique gene configurations, extending the effective range to $[0, 1+\delta_{\max}] \approx [0, 2.8]$. Values above 1.0 reported in Table~2 reflect this combined $F' = F + \delta$ score.
    \item \textbf{Convergence detection}: $\max(f_{t-9:t}) - \min(f_{t-9:t}) < \epsilon$.
\end{itemize}

\subsection{GPU Batch Evaluation}

The \texttt{GPUBatchEvaluator} computes all four fitness metrics as batched tensor operations on GPU, processing 1,000 genes in a single kernel launch. Genes are padded to uniform length and stacked into a $(B \times L)$ tensor.

\subsection{HTCV4 Population Compression}

The \texttt{PopulationCompressor} serializes entire populations via HTCV4: all gene weights are concatenated, compressed with block deduplication + trit packing + ZSTD, and stored alongside JSON metadata preserving gene IDs, fitness scores, and lineage.

% ==================== EXPERIMENTS ====================
\section{Experiments}

\subsection{Setup}

\begin{itemize}
    \item \textbf{Gene pool}: 200 synthetic genes, lengths $\in \{768, 1024, 3072, 4096\}$ (GPT-2 transformer dimensions), distribution $(-1: 29\%, 0: 42\%, +1: 29\%)$.
    \item \textbf{Hardware}: AMD RX 6600M (8\,GB VRAM), ROCm 6.2, PyTorch 2.5.1.
    \item \textbf{Software}: Python 3.12.3, NumPy, seed=42 for reproducibility.
\end{itemize}

\subsection{Results}

\begin{table}[H]
\centering
\caption{Single-pass synthesis pipeline (500 candidates)}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{Unit} \\
\midrule
Candidates generated & 550 & genes \\
Top-50 selected & 50 & genes \\
Best fitness & 0.9267 & score \\
Average fitness (top-50) & 0.9263 & score \\
Pipeline time & 427.8 & ms \\
\midrule
\multicolumn{3}{@{}l}{\textit{By synthesis method:}} \\
\quad Mutation & 150 & genes \\
\quad Interpolation & 100 & genes \\
\quad Crossover & 100 & genes \\
\quad Distribution sample & 100 & genes \\
\quad $\varphi$-optimized & 100 & genes \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Evolutionary optimization (30 generations)}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{Unit} \\
\midrule
Generations run & 30 & gens \\
Best fitness & 2.5419 & score \\
Average fitness & 2.1107 & score \\
Final mutation rate & 0.0100 & rate \\
Evolution time & 0.83 & s \\
\midrule
\multicolumn{3}{@{}l}{\textit{Diversity metrics:}} \\
\quad Genetic entropy & 0.9884 & norm. \\
\quad Effective $N_e$ & 96.5 & genes \\
\quad Unique patterns & 3,296 & count \\
\quad Pairwise distance & 0.6396 & Hamming \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{GPU evaluation and compression}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{Unit} \\
\midrule
GPU batch eval (1000 genes) & 121.5 & ms \\
Throughput & 8,229 & genes/s \\
Device & \multicolumn{2}{l}{AMD RX 6600M (HIP)} \\
\midrule
Population elements & 366,292 & trits \\
HTCV4 compressed & 40,667 & bytes \\
Compression ratio & 9.0 & :1 \\
Roundtrip verified & \multicolumn{2}{l}{100 genes $\checkmark$} \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
    width=\columnwidth,
    height=4.5cm,
    xlabel={Generation},
    ylabel={Best Fitness},
    grid=major,
    mark size=1.2pt,
    thick,
    title={\scriptsize Evolutionary Convergence (30 generations)}
]
\addplot[arkblue, mark=*] coordinates {
    (0,0.89) (3,1.15) (6,1.48) (9,1.72) (12,1.95)
    (15,2.12) (18,2.28) (21,2.38) (24,2.45) (27,2.50) (29,2.54)
};
\addplot[arkred, mark=square*, dashed] coordinates {
    (0,0.10) (3,0.10) (6,0.08) (9,0.06) (12,0.04)
    (15,0.03) (18,0.02) (21,0.015) (24,0.012) (27,0.010) (29,0.010)
};
\legend{\scriptsize Best Fitness, \scriptsize Mutation Rate}
\end{axis}
\end{tikzpicture}
\caption{Best fitness and adaptive mutation rate over 30 generations. Mutation rate decreases as fitness improves.}
\end{figure}

\subsection{HUAM Memory Integration}

High-fitness genes ($F > 0.5$) are persisted to HUAM hierarchical memory, enabling cross-session reuse. On subsequent runs, \texttt{load\_genes\_from\_huam()} retrieves previously evolved genes, seeding the population with proven individuals and reducing convergence time.

% ==================== DISCUSSION ====================
\section{Discussion}

\subsection{Fitness Beyond Proxy}

The proxy fitness (structural metrics only) correlates with model performance but is not a substitute for full training evaluation. The adaptive combination with $\varphi$-weights provides a useful heuristic, but real downstream task evaluation remains necessary for production deployment.

\noindent\textbf{Validation Gap:} The proxy fitness metrics (sparsity, balance, coherence, entropy) have not been validated against downstream model performance (perplexity, accuracy). Establishing this correlation is essential future work.

\subsection{Scalability}

At 8,229 genes/second on GPU, evaluating $10^6$ candidate genes requires $\sim$2 minutes. Population compression at 9:1 enables storing $10^4$-gene populations in $<$5\,MB. The system scales to LLM-size gene pools (millions of genes) through sampling-based analysis and batched evaluation.

\subsection{Limitations}

\begin{enumerate}
    \item Proxy fitness does not capture task-specific performance.
    \item Single-objective optimization; multi-objective Pareto fronts are unexplored.
    \item No formal convergence guarantees for evolutionary search.
    \item $\varphi$-optimization is a heuristic, not proven optimal.
\end{enumerate}

% ==================== RELATED WORK ====================
\section{Related Work}

Binary and ternary quantization~\cite{li2016ternary,rastegari2016xnor} focus on inference efficiency. Neuroevolution~\cite{stanley2019designing} typically operates on continuous weights or architectures. Our work uniquely combines \textbf{discrete parameter evolution} with \textbf{trit-native compression} and \textbf{proxy fitness evaluation}.

% ==================== CONCLUSION ====================
\section{Conclusion}

We presented Gene Synthesis, a complete pipeline for generating, evaluating, and evolving ternary neural network parameters. The discrete nature of $\{-1,0,+1\}$ weights enables exact analysis, meaningful recombination, and lossless population compression---capabilities impossible in continuous parameter spaces. Our implementation achieves sub-second synthesis (427.8\,ms for 500 candidates), GPU-accelerated evaluation (8,229 genes/s), and evolutionary convergence in 30 generations (0.83\,s). The system's 2,536 LOC implementation passes 41/41 unit tests and integrates with the ARKHEION Nucleus via HTCV4 compression and HUAM memory persistence.\footnote{Implementation update (Feb 2026): The gene synthesis ecosystem has since expanded to 27 Python source files (~19K LOC), incorporating additional evolution strategies, population management, and extended HUAM integration. The 2,536 SLOC figure reflects the core pipeline described in this paper.}

% ==================== REFERENCES ====================
\begin{thebibliography}{9}
\bibitem{li2016ternary} F.~Li et al., ``Ternary Weight Networks,'' \textit{arXiv:1605.04711}, 2016.
\bibitem{rastegari2016xnor} M.~Rastegari et al., ``XNOR-Net: ImageNet Classification Using Binary Convolutional Neural Networks,'' \textit{ECCV}, 2016.
\bibitem{stanley2019designing} K.~O.~Stanley et al., ``Designing neural networks through neuroevolution,'' \textit{Nature Machine Intelligence}, 1(1):24--35, 2019.
\bibitem{vogel1979better} H.~Vogel, ``A better way to construct the sunflower head,'' \textit{Mathematical Biosciences}, 44:179--189, 1979.
\end{thebibliography}

\end{document}
